PANCHANGE_data_preprocessing

### 2020-08-13 PANCHANGE_data_preprocessing2020-07-15.Rmd
Amended psychosis sections. Removed dplyr from packages because it's part of tidyverse.

#Packages
Install packages (if they are not available in your version of R)
```{r Installing packages}
#install.packages("summarytools")
#install.packages("tidyverse")
#install.packages("psych")
#install.packages("dplyr")
#install.packages("eeptools")
```

Load packages
```{r Load packages}
library(knitr)
library(summarytools)
library(psych)
library(dplyr)
library(tidyverse)
library(eeptools)
```

@@

```{r nbr cohort recode ethnicity}
#Recode details ethnicity info
nbr.cohort.raw.id$Ethnicity_full <- recode_factor(nbr.cohort.raw.id$Ethnicity.unc,
                                             "A" = "White - British",
                                             "B" = "White - Irish",
                                             "C" = "White - Other",
                                             "D" = "Mixed - White & Black Caribbean",
                                             "E" = "Mixed - White & Black African",
                                             "F" = "Mixed - White & Asian",
                                             "G" = "Mixed - Other",
                                             "H" = "Asian - Indian",
                                             "J" = "Asian - Pakistani",
                                             "K" = "Asian - Bangladeshi",
                                             "L" = "Asian - Other",
                                             "M" = "Black - Caribbean",
                                             "N" = "Black - African",
                                             "P" = "Black - Other",
                                             "R" = "Chinese - Chinese",
                                             "S" = "Chinese - Other",
                                             "W" = NULL, ## Asking NBR what this level means
                                             "Z" = NULL,
                                             "99" = NULL)

#Match GLAD/EDGI/RAMP categories
nbr.cohort.raw.id$Ethnicity <- recode_factor(nbr.cohort.raw.id$Ethnicity.unc,
                                             "A" = "White, white European or Caucasian",
                                             "B" = "White, white European or Caucasian",
                                             "C" = "White, white European or Caucasian",
                                             "D" = "Mixed or multiple ethnic origins",
                                             "E" = "Mixed or multiple ethnic origins",
                                             "F" = "Mixed or multiple ethnic origins",
                                             "G" = "Mixed or multiple ethnic origins",
                                             "H" = "Asian or Asian British (including Chinese)",
                                             "J" = "Asian or Asian British (including Chinese)",
                                             "K" = "Asian or Asian British (including Chinese)",
                                             "L" = "Asian or Asian British (including Chinese)",
                                             "M" = "Black or Black British",
                                             "N" = "Black or Black British",
                                             "P" = "Black or Black British",
                                             "R" = "Asian or Asian British (including Chinese)",
                                             "S" = "Asian or Asian British (including Chinese)",
                                             "W" = NULL, ## Asking NBR what this level means
                                             "Z" = NULL,
                                             "99" = NULL)

@@

# Categorise age into groups 
In RAMP, age is indicated categorically: 16-18, 19-25, 26-35, 36-45, 46-55, 56-65, 66-70, 71-75, 76-80, 81-85, 86-90, 91-100, 100+
Age categories in GLAD, EDGI and NBR have been created to reflect this
```{r Age groups}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE ) {

# Create categorical age groups per 10 years
dem.raw.id <- dem.raw.id %>%
  mutate(
    age_category_numeric =
      case_when(
          Age >= 16 & Age <= 18 ~ "1",
          Age >= 19 & Age <= 25 ~ "2",
          Age >= 26 & Age <= 35 ~ "3",
          Age >= 36 & Age <= 45 ~ "4",
          Age >= 46 & Age <= 55 ~ "5",
          Age >= 56 & Age <= 65 ~ "6",
          Age >= 66 & Age <= 70 ~ "7",
          Age >= 71 & Age <= 75 ~ "8",
          Age >= 76 & Age <= 80 ~ "9",
          Age >= 81 & Age <= 85 ~ "10",
          Age >= 86 & Age <= 90 ~ "11",
          Age >= 91 & Age <= 100 ~ "12",
          Age >= 101 & Age <= 120 ~ "13" # oldest person in the world is 117 years
                )
      )

@@

# Rename columns that differ from GLAD
nbr.mhd.raw.id <- nbr.mhd.raw.id %>% 
  mutate(
    mhd.specific_phobia_e.g._phobia_of_flying = mhd.specific_phobia_e.g._fear_of_flying,
    mhd.specific_phobia_e.g._phobia_of_flying_numeric = mhd.specific_phobia_e.g._fear_of_flying_numeric,
    mhd.psychological_overeating_or_bingeeating = mhd.bingeeating_disorder,
    mhd.psychological_overeating_or_bingeeating_numeric = mhd.bingeeating_disorder_numeric
  )

# Inspect dimensions
dim(nbr.mhd.raw.id)
# Inspect colnames
colnames(nbr.mhd.raw.id)
#Differences
dim(nbr.mhd.raw)[1]-dim(nbr.mhd.raw.id)[1]
}
```

```{r nbr mhd copy}
if(NBR == TRUE) {
  mhd.raw.id <- nbr.mhd.raw.id
}
```

@@

Obsessive compulsive disorders
```{r glad mhd obsessive_compulsive_disorders}
mhd.raw.id <- mhd.raw.id %>%
  mutate(
    obsessive_compulsive_disorders_numeric =
      case_when(
        mhd.obsessivecompulsive_disorder_ocd_numeric == "1" ~ 1,
        mhd.obsessive_compulsive_related_disorders_numeric == "1" ~ 1,
        mhd.body_dysmorphic_disorder_bdd_numeric == "1" ~ 1,
        (mhd.obsessivecompulsive_disorder_ocd_numeric == "0" &
        mhd.obsessive_compulsive_related_disorders_numeric == "0" &
        mhd.body_dysmorphic_disorder_bdd_numeric == "0") ~ 0
      )
  )

@@

mhd.raw.id %>%
  freq(obsessive_compulsive_disorders)
```

```{r Check obsessive_compulsive_disorders}
mhd.raw.id %>%
  select(
    obsessive_compulsive_disorders,
    obsessive_compulsive_disorders_numeric,
    mhd.obsessivecompulsive_disorder_ocd,
    mhd.obsessivecompulsive_disorder_ocd_numeric,
    mhd.obsessive_compulsive_related_disorders,
    mhd.obsessive_compulsive_related_disorders_numeric
  )
```

Psychotic disorders group
[63] "mhd.schizophrenia_numeric"                                 
[64] "mhd.schizoaffective_disorder_numeric"                      
[65] "mhd.type_psychosis_psychotic_illness_numeric" 
```{r glad mhd psychotic_disorders}
mhd.raw.id <- mhd.raw.id %>%
  mutate(
    psychotic_disorders_numeric =
      case_when(
        mhd.schizophrenia_numeric == "1" ~ 1,
        mhd.schizoaffective_disorder_numeric == "1" ~ 1,
        mhd.type_psychosis_psychotic_illness_numeric == "1" ~ 1,
        (mhd.schizophrenia_numeric == "0" &
        mhd.schizoaffective_disorder_numeric == "0" &
        mhd.type_psychosis_psychotic_illness_numeric == "0") ~ 0
      )
  )

@@

mhd.raw.id %>%
  freq(psychotic_disorders)
```

Overlap between self-report schizophrenia and bipolar disorder
+++CH: Check overlap between schizophrenia group and bipolar disorder
++KT: Only 63 with comorbid self-report bipolar disorder and schizophrenia. More likely to have shared symptoms, less likely to have confirmed diagnoses of both. Could be too severe to be in GLAD?
```{r glad mhd bipolar_and_schizophrenia_numeric}
mhd.raw.id <- mhd.raw.id %>%
  mutate(
    bipolar_and_schizophrenia_numeric =
      case_when(
        psychotic_disorders_numeric == 0 &
          mhd.mania_hypomania_bipolar_or_manicdepression_numeric == 0 ~ 0,
                psychotic_disorders_numeric == 1 &
          mhd.mania_hypomania_bipolar_or_manicdepression_numeric == 0 ~ 2,
        psychotic_disorders_numeric == 0 &
          mhd.mania_hypomania_bipolar_or_manicdepression_numeric == 1 ~ 3,
                psychotic_disorders_numeric == 1 &
          mhd.mania_hypomania_bipolar_or_manicdepression_numeric == 1 ~ 1
      )
  )

@@

mhd.raw.id %>%
  freq(bipolar_and_schizophrenia,
       cumul = F)
```

```{r Check psychotic_disorders}
mhd.raw.id %>%
  select(
    psychotic_disorders,
    psychotic_disorders_numeric,
    bipolar_and_schizophrenia,
    bipolar_and_schizophrenia_numeric,
    mhd.schizophrenia_numeric,
    mhd.schizoaffective_disorder_numeric,
    mhd.mania_hypomania_bipolar_or_manicdepression,
    mhd.mania_hypomania_bipolar_or_manicdepression_numeric,
    mhd.type_psychosis_psychotic_illness_numeric,
    mhd.schizophrenia,
    mhd.schizoaffective_disorder,
    mhd.type_psychosis_psychotic_illness
  )
```

>>>>

#Packages
Install packages (if they are not available in your version of R)
```{r Installing packages}
#install.packages("summarytools")
#install.packages("tidyverse")
#install.packages("psych")
#install.packages("eeptools")
```

Load packages
```{r Load packages}
library(knitr)
library(summarytools)
library(psych)
library(tidyverse)
library(eeptools) #Needed to calculate age from birthdate
```

@@

```{r nbr cohort recode ethnicity}
if(NBR == TRUE) {
#Recode details ethnicity info
nbr.cohort.raw.id$Ethnicity_full <- recode_factor(nbr.cohort.raw.id$Ethnicity.unc,
                                             "A" = "White - British",
                                             "B" = "White - Irish",
                                             "C" = "White - Other",
                                             "D" = "Mixed - White & Black Caribbean",
                                             "E" = "Mixed - White & Black African",
                                             "F" = "Mixed - White & Asian",
                                             "G" = "Mixed - Other",
                                             "H" = "Asian - Indian",
                                             "J" = "Asian - Pakistani",
                                             "K" = "Asian - Bangladeshi",
                                             "L" = "Asian - Other",
                                             "M" = "Black - Caribbean",
                                             "N" = "Black - African",
                                             "P" = "Black - Other",
                                             "R" = "Chinese - Chinese",
                                             "S" = "Chinese - Other",
                                             "W" = NULL, ## Asking NBR what this level means
                                             "Z" = NULL,
                                             "99" = NULL)

#Match GLAD/EDGI/RAMP categories
nbr.cohort.raw.id$Ethnicity <- recode_factor(nbr.cohort.raw.id$Ethnicity.unc,
                                             "A" = "White, white European or Caucasian",
                                             "B" = "White, white European or Caucasian",
                                             "C" = "White, white European or Caucasian",
                                             "D" = "Mixed or multiple ethnic origins",
                                             "E" = "Mixed or multiple ethnic origins",
                                             "F" = "Mixed or multiple ethnic origins",
                                             "G" = "Mixed or multiple ethnic origins",
                                             "H" = "Asian or Asian British (including Chinese)",
                                             "J" = "Asian or Asian British (including Chinese)",
                                             "K" = "Asian or Asian British (including Chinese)",
                                             "L" = "Asian or Asian British (including Chinese)",
                                             "M" = "Black or Black British",
                                             "N" = "Black or Black British",
                                             "P" = "Black or Black British",
                                             "R" = "Asian or Asian British (including Chinese)",
                                             "S" = "Asian or Asian British (including Chinese)",
                                             "W" = NULL, ## Asking NBR what this level means
                                             "Z" = NULL,
                                             "99" = NULL)
}
```

@@

# Categorise age into groups 
In RAMP, age is indicated categorically: 16-18, 19-25, 26-35, 36-45, 46-55, 56-65, 66-70, 71-75, 76-80, 81-85, 86-90, 91-100, 100+
Age categories in GLAD, EDGI and NBR have been created to reflect this
```{r Age groups}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE ) {
# Create categorical age groups per 10 years
dem.raw.id <- dem.raw.id %>%
  mutate(
    age_category_numeric =
      case_when(
          Age >= 16 & Age <= 18 ~ "1",
          Age >= 19 & Age <= 25 ~ "2",
          Age >= 26 & Age <= 35 ~ "3",
          Age >= 36 & Age <= 45 ~ "4",
          Age >= 46 & Age <= 55 ~ "5",
          Age >= 56 & Age <= 65 ~ "6",
          Age >= 66 & Age <= 70 ~ "7",
          Age >= 71 & Age <= 75 ~ "8",
          Age >= 76 & Age <= 80 ~ "9",
          Age >= 81 & Age <= 85 ~ "10",
          Age >= 86 & Age <= 90 ~ "11",
          Age >= 91 & Age <= 100 ~ "12",
          Age >= 101 & Age <= 120 ~ "13" # oldest person in the world is 117 years
                )
      )

@@

# Rename columns that differ from GLAD
nbr.mhd.raw.id_renamed <- nbr.mhd.raw.id %>% 
  mutate(
    mhd.specific_phobia_e.g._phobia_of_flying = mhd.specific_phobia_e.g._fear_of_flying,
    mhd.specific_phobia_e.g._phobia_of_flying_numeric = mhd.specific_phobia_e.g._fear_of_flying_numeric,
    mhd.psychological_overeating_or_bingeeating = mhd.bingeeating_disorder,
    mhd.psychological_overeating_or_bingeeating_numeric = mhd.bingeeating_disorder_numeric,
    mhd.type_psychosis_psychotic_illness = mhd.psychosis_type_psychotic_illness,
    mhd.type_psychosis_psychotic_illness_numeric = mhd.psychosis_type_psychotic_illness_numeric
    ) %>% 
  add_column(
    mhd.schizoaffective_disorder = NA, #NBR dooesn't ask about schizoaffective
    mhd.schizoaffective_disorder_numeric = NA
  )

# Inspect dimensions
dim(nbr.mhd.raw.id_renamed)
# Inspect colnames
colnames(nbr.mhd.raw.id_renamed)
#Differences
dim(nbr.mhd.raw)[1]-dim(nbr.mhd.raw.id_renamed)[1]
}
```

```{r nbr mhd copy}
if(NBR == TRUE) {
  mhd.raw.id <- nbr.mhd.raw.id_renamed
}
```

@@

Obsessive compulsive disorders
```{r mhd obsessive_compulsive_disorders}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE | RAMP == TRUE) {
mhd.raw.id <- mhd.raw.id %>%
  mutate(
    obsessive_compulsive_disorders_numeric =
      case_when(
        mhd.obsessivecompulsive_disorder_ocd_numeric == "1" ~ 1,
        mhd.obsessive_compulsive_related_disorders_numeric == "1" ~ 1,
        mhd.body_dysmorphic_disorder_bdd_numeric == "1" ~ 1,
        (mhd.obsessivecompulsive_disorder_ocd_numeric == "0" &
        mhd.obsessive_compulsive_related_disorders_numeric == "0" &
        mhd.body_dysmorphic_disorder_bdd_numeric == "0") ~ 0
      )
  )

@@

mhd.raw.id %>%
  freq(obsessive_compulsive_disorders)
}
```

```{r Check obsessive_compulsive_disorders}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE | RAMP == TRUE) {
mhd.raw.id %>%
  select(
    obsessive_compulsive_disorders,
    obsessive_compulsive_disorders_numeric,
    mhd.obsessivecompulsive_disorder_ocd,
    mhd.obsessivecompulsive_disorder_ocd_numeric,
    mhd.obsessive_compulsive_related_disorders,
    mhd.obsessive_compulsive_related_disorders_numeric
  )
}
```

Psychotic disorders group
[63] "mhd.schizophrenia_numeric"                                 
[64] "mhd.schizoaffective_disorder_numeric"                      
[65] "mhd.type_psychosis_psychotic_illness_numeric" 
```{r mhd psychotic_disorders}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE | RAMP == TRUE) {
mhd.raw.id <- mhd.raw.id %>%
  mutate(
    psychotic_disorders_numeric =
      case_when(
        mhd.schizophrenia_numeric == "1" ~ 1,
        mhd.schizoaffective_disorder_numeric == "1" ~ 1,
        mhd.type_psychosis_psychotic_illness_numeric == "1" ~ 1,
        (mhd.schizophrenia_numeric == "0" &
        (mhd.schizoaffective_disorder_numeric == "0" | is.na(mhd.schizoaffective_disorder_numeric == "0") &
        mhd.type_psychosis_psychotic_illness_numeric == "0") ~ 0
      )
  )

@@

mhd.raw.id %>%
  freq(psychotic_disorders)
}
```

Overlap between self-report schizophrenia and bipolar disorder
+++CH: Check overlap between schizophrenia group and bipolar disorder
++KT: Only 63 with comorbid self-report bipolar disorder and schizophrenia. More likely to have shared symptoms, less likely to have confirmed diagnoses of both. Could be too severe to be in GLAD?
```{r mhd bipolar_and_schizophrenia_numeric}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE | RAMP == TRUE) {
mhd.raw.id <- mhd.raw.id %>%
  mutate(
    bipolar_and_schizophrenia_numeric =
      case_when(
        psychotic_disorders_numeric == 0 &
          mhd.mania_hypomania_bipolar_or_manicdepression_numeric == 0 ~ 0,
                psychotic_disorders_numeric == 1 &
          mhd.mania_hypomania_bipolar_or_manicdepression_numeric == 0 ~ 2,
        psychotic_disorders_numeric == 0 &
          mhd.mania_hypomania_bipolar_or_manicdepression_numeric == 1 ~ 3,
                psychotic_disorders_numeric == 1 &
          mhd.mania_hypomania_bipolar_or_manicdepression_numeric == 1 ~ 1
      )
  )

@@

mhd.raw.id %>%
  freq(bipolar_and_schizophrenia,
       cumul = F)
}
```

```{r Check psychotic_disorders}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE | RAMP == TRUE) {
mhd.raw.id %>%
  select(
    psychotic_disorders,
    psychotic_disorders_numeric,
    bipolar_and_schizophrenia,
    bipolar_and_schizophrenia_numeric,
    mhd.schizophrenia_numeric,
    mhd.schizoaffective_disorder_numeric,
    mhd.mania_hypomania_bipolar_or_manicdepression,
    mhd.mania_hypomania_bipolar_or_manicdepression_numeric,
    mhd.type_psychosis_psychotic_illness_numeric,
    mhd.schizophrenia,
    mhd.schizoaffective_disorder,
    mhd.type_psychosis_psychotic_illness
  )
}
```
